\LARGE \bf Towards a Data-driven Approach to Motion Planning

### 2015-02-24 paper2015asml.tex
latest changes

\overrideIEEEmargins                                      % Needed to meet printer requirements.

@@

This paper is organized as follows. In Section~\ref{sec:background}, we first describe related work in this area. In Section~\ref{sec:survey}, we present details of the online survey including information about the types of trajectories and environments used. Section~\ref{sec:raw_results} presents the results from the survey and analysis, including the trajectory features used for classification and Section~\ref{sec:classification} covers the classification approaches used with the data. In Section~\ref{sec:conclusion}, we conclude with a discussion of how our results could be used to design motion planners that generate more human-acceptable trajectories. 
\begin{figure}
\centering
\begin{tabular}{c|c}
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}} &
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}} \\
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}} &
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}} \\
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}} &
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}} \\
(a) & (b) \\
\end{tabular}
%\begin{minipage}[][][s]{0.45\columnwidth}
%\centering
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{2}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{4}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}}
%(a)
%\end{minipage}
%\hfill
%\vline
%\hfill
%\begin{minipage}[][][s]{0.45\columnwidth}
%\centering
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{2}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{4}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}}
%(b)
%\end{minipage}
\caption{Example trajectory snapshots for motions that survey participants rated (a) highly and (b) poorly. In (b) the hand of the robot ducks under the table in between the first and middle snapshot.}\label{fig:example_trajectory}
\end{figure}

@@

\begin{table*}[t]
\centering
\scalebox{0.9}{
\begin{tabular}{cccccccc}
&&&&&&&\\
\hline
Efficiency  & Highly inefficient& Moderately inefficient& Slightly inefficient& Neutral & Slightly efficient& Moderately efficient& Highly efficient\\
\hline
Elegance    & Highly awkward& Moderately awkward& Slightly awkward& Neutral & Slightly elegant& Moderately elegant& Highly elegant\\
\hline
Smoothness  & Highly rough& Moderately rough& Slightly rough& Neutral & Slightly smooth& Moderately smooth& Highly smooth\\
\hline
Overall     & Extremely poor& Moderately poor& Somewhat poor& Neutral & Somewhat good& Moderately good& Extremely good\\
\hline
\end{tabular}
}
\caption{The four rating categories/attributes and the associated scales for each presented to the survey respondent.}
\label{tab:rating_scales}
\end{table*}

@@

\section{Survey Results and Analysis}
\label{sec:raw_results}
The survey resulted in at least 20 ratings for each trajectory (so there were a total of 2060 ratings from participants in the survey). In this section, we will first analyze the raw results from the survey (which are interesting in themselves). Subsequently, we will present a set of features that were used to represent individual trajectories. These features will be used in the next section to learn the human-observer ratings for robot trajectories. 
\subsection{Raw Survey Results}
\begin{figure}[h]
  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
  \caption{Histogram of averaged ratings for the 4 different attributes for 103 trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Awkward, Smooth vs. Rough, Overall Rating}
\label{fig:survey_raw}
\end{figure}
The raw results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of averaged responses for each of the four survey questions across the 103 trajectories that were rated by users. It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well. An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall). Table~\ref{tab:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes. Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$). The 95\% confidence intervals for all the four ratings for each trajectory was less than 1.0. 
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
& Efficient & Elegant & Smooth & Overall \\ \hline
Efficient & 1 & 0.96 & 0.91 & 0.98 \\ \hline
Elegant & 0.96 & 1 & 0.96 & 0.98 \\ \hline
Smooth & 0.91 & 0.96 & 1 & 0.96 \\ \hline
Overall & 0.98 & 0.98 & 0.96 & 1 \\ \hline
\end{tabular}
\caption{Correlation between the four attributes in the survey.}
\label{tab:correlation}
\vspace{-0.1in}
\end{table}

@@

The classification results clearly show the value in using a data-driven approach to determining human preferences for robot trajectories. It is difficult, 
in obstacle filled cluttered environments, to hand-design an appropriate cost function for generating human preferred trajectories. Tuning coefficients 
and weights for all the individual cost components for designing preferred trajectories requires significant effort. 
The classifiers that we have built give us a notion of the quality of trajectories 
being generated without an explicit need to model the cost function. The output from the classifier could be integrated into gradient-free motion planning schemes 
(e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
should also help in the design of cost functions, when necessary, especially since some the features may not be traditionally included as part of a cost function. 
In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which mostly relies on a path length optimization criterion.

\section{Conclusion}
\label{sec:conclusion}
In this work, we have gathered ratings for a large dataset of robot trajectories. We have already gained significant insights into the 
types of trajectories that are preferred by human observers. In particular, we have been able to identify a set of features that seem to influence 
the human's perception of the trajectories the most. This can help us in defining better cost functions when generating the trajectories. The classifiers 
that we have learned could also be used to directly inform gradient free motion planning algorithms by assigning a rating to new trajectories. Our 
results were generated using a single robot model (the PR2) but we believe that they will easily translate to other robot models as well. To this end, 
we will keep the survey open and attempt to collect data for a larger set of trajectories in 
more environments and with different robots.

>>>>

%\overrideIEEEmargins                                      % Needed to meet printer requirements.

@@

This paper is organized as follows. In Section~\ref{sec:background}, we first describe related work in this area. In Section~\ref{sec:survey}, we present details of the online survey including information about the types of trajectories and environments used. Section~\ref{sec:raw_results} presents the results from the survey and analysis, including the trajectory features used for classification and Section~\ref{sec:classification} covers the classification approaches used with the data. In Section~\ref{sec:conclusion}, we conclude with a discussion of how our results could be used to design motion planners that generate more human-acceptable trajectories. 
\begin{figure}
\centering
\begin{tabular}{c|c}
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}} &
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}} \\
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}} &
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}} \\
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}} &
\adjincludegraphics[trim = 100mm 0mm 100mm 30mm, width=0.44\columnwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}} \\
(a) & (b) \\
\end{tabular}
%\begin{minipage}[][][s]{0.45\columnwidth}
%\centering
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{1}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{2}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{3}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{4}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\gta{2014_Sep_17_18_40_15}{2}{5}}
%(a)
%\end{minipage}
%\hfill
%\vline
%\hfill
%\begin{minipage}[][][s]{0.45\columnwidth}
%\centering
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{1}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{2}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{3}}
%%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{4}}
%\includegraphics[trim = 100mm 0mm 100mm 30mm, width=\textwidth, clip=true]{\bta{2014_Sep_17_18_40_21}{2}{5}}
%(b)
%\end{minipage}
\caption{Example trajectory snapshots for motions that survey participants rated (a) highly and (b) poorly. In (b) the hand of the robot ducks under the table in between the first and middle snapshot.}\label{fig:example_trajectory}
\vspace{-0.2in}
\end{figure}

@@

\begin{table*}[t]
\centering
\scalebox{0.9}{
\begin{tabular}{cccccccc}
&&&&&&&\\
\hline
Efficiency  & Highly inefficient& Moderately inefficient& Slightly inefficient& Neutral & Slightly efficient& Moderately efficient& Highly efficient\\
\hline
Elegance    & Highly awkward& Moderately awkward& Slightly awkward& Neutral & Slightly elegant& Moderately elegant& Highly elegant\\
\hline
Smoothness  & Highly rough& Moderately rough& Slightly rough& Neutral & Slightly smooth& Moderately smooth& Highly smooth\\
\hline
Overall     & Extremely poor& Moderately poor& Somewhat poor& Neutral & Somewhat good& Moderately good& Extremely good\\
\hline
\end{tabular}
}
\caption{The four rating categories/attributes and the associated scales for each presented to the survey respondent.}
\label{tab:rating_scales}
\vspace{-0.2in}
\end{table*}

@@

\section{Survey Results and Analysis}
\label{sec:raw_results}
The survey resulted in at least 20 ratings for each trajectory (so there were a total of 2060 ratings from participants in the survey). In this section, we will first analyze the raw results from the survey (which are interesting in themselves). Subsequently, we will present a set of features that were used to represent individual trajectories. These features will be used in the next section to learn the human-observer ratings for robot trajectories. 
\subsection{Raw Survey Results}
\begin{figure}[h]
  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
  \caption{Histogram of averaged ratings for the 4 different attributes for 103 trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Awkward, Smooth vs. Rough, Overall Rating}
\label{fig:survey_raw}
\end{figure}
The raw results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of averaged responses for each of the four survey questions across the 103 trajectories that were rated by users. It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well. An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall). Table~\ref{tab:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes. Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$). The Spearman rank correlation coefficients are also high and statistically significant. 
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
& Efficient & Elegant & Smooth & Overall \\ \hline
Efficient & 1 & 0.96 & 0.91 & 0.98 \\ \hline
Elegant & 0.96 & 1 & 0.96 & 0.98 \\ \hline
Smooth & 0.91 & 0.96 & 1 & 0.96 \\ \hline
Overall & 0.98 & 0.98 & 0.96 & 1 \\ \hline
\end{tabular}
\caption{Correlation between the four attributes in the survey.}
\label{tab:correlation}
\vspace{-0.1in}
\end{table}

@@

The classification results clearly show the value in using a data-driven approach to determining human preferences for robot trajectories. It is difficult, 
in obstacle filled cluttered environments, to hand-design an appropriate cost function for generating human preferred trajectories. Tuning coefficients 
and weights for all the individual cost components for designing preferred trajectories requires significant effort. 
The classifiers that we have built give us a notion of the quality of trajectories 
being generated without an explicit need to model the cost function. The output from the classifier could be integrated into gradient-free motion planning schemes 
(e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
should also help in the design of cost functions, especially since some the features may not be traditionally included as part of a cost function. 
In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which rely on a path length criterion.

\section{Conclusion}
\label{sec:conclusion}
In this work, we have gathered ratings for a large dataset of robot trajectories. We have already gained significant insights into the 
types of trajectories that are preferred by human observers. In particular, we have been able to identify a set of features that seem to influence 
the human's perception of the trajectories the most. This can help us in defining better cost functions when generating the trajectories. The classifiers 
that we have learned could also be used to directly inform gradient free motion planning algorithms by assigning a rating to new trajectories. Our 
results were generated using the PR2 but we believe that they will easily translate to other robot models as well. To this end, 
we will keep the survey open and attempt to collect data for a larger set of trajectories in 
more environments and with different robots.
